---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)

toc <- function() {
  re <- readLines("README.Rmd")
  has_title <- as.logical(lapply(re, function(x) substr(x, 1, 2) == "##"))
  only_titles <- re[has_title]
  titles <- trimws(gsub("#", "", only_titles))
  links <- trimws(gsub("`", "", titles))
  links <- tolower(links)
  links <- trimws(gsub(" ", "-", links))
  toc_list <- lapply(
    seq_along(titles),
    function(x) {
      if(substr(only_titles[x], 1, 3) == "###") {
        paste0("<ul><li><a href=#", links[x], ">", titles[x], "</a></li></ul>")  
      } else {
        paste0("<li><a href=#", links[x], ">", titles[x], "</a></li>")  
      }
    }
  )
  toc_full <- paste(toc_list, collapse = "") 
  paste0("<ul>", toc_full, "</ul>")
}
```

# connections

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/edgararuiz/connections.svg?branch=master)](https://travis-ci.com/edgararuiz/connections)
[![Codecov test coverage](https://codecov.io/gh/edgararuiz/connections/branch/master/graph/badge.svg)](https://codecov.io/gh/edgararuiz/connections?branch=master)
<!-- badges: end -->

`r toc()`

Provides a generic implementation of the [RStudio Connection Contract](https://rstudio.github.io/rstudio-extensions/connections-contract.html) to make it easier for database connections, and other type of connections, opened via R packages to take advantage of the Connections Pane inside the RStudio IDE.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("edgararuiz/connections")
```

## Usage

### Basic

```{r}
library(connections)
library(RSQLite)
library(DBI)

con <- connection_open(SQLite(), path = ":dbname")
```

<img src = "man/figures/sqlite-1.png" width = "400px"> <br/>

```{r}
dbWriteTable(con, "db_mtcars", mtcars)
```

<img src = "man/figures/sqlite-2.png" width = "400px"> <br/>

Preview button works as expected

<img src = "man/figures/sqlite-3.png" width = "600px"> <br/>

```{r}
connection_close(con)
```

<img src = "man/figures/sqlite-4.png" width = "400px"> <br/>

### Not integrated

```{r}
con <- dbConnect(SQLite(), path = ":dbname")
```



```{r}
connection_view(con)
```

<img src = "man/figures/sqlite-5.png" width = "400px"> <br/>

```{r}
dbWriteTable(con, "db_mtcars", mtcars)
```

```{r}
connection_update(con)
```

<img src = "man/figures/sqlite-6.png" width = "400px"> <br/>

```{r}
connection_close(con)
```

<img src = "man/figures/sqlite-7.png" width = "400px"> <br/>

### Pre-set names

```{r}
con <- dbConnect(SQLite(), path = ":dbname")
```


```{r}
connection_view(
  con, 
  host = "my_host", 
  name = "my_name",
  connection_code = "library(connections)\ndbConnect(...)"
  )
```


<img src = "man/figures/sqlite-8.png" width = "400px"> <br/>

Connection code is sourced from `connection_code`

<img src = "man/figures/sqlite-9.png" width = "400px"> <br/>

## `DBI` packages examples

```{r}
library(RPostgres)
library(connections)
library(DBI)

con <- connection_open(Postgres(), 
                 host = "sol-eng-postgre.cihykudhzbgw.us-west-2.rds.amazonaws.com",
                 dbname = "finance",
                 user = "xxxxx",
                 password = "xxxxx",
                 port = 5432
                 )

```

<img src = "man/figures/postgres-1.png" width = "400px"> <br/>

<img src = "man/figures/postgres-2.png" width = "400px"> <br/>

## Custom connections

```{r}
library(connections)

my_conn <-  list(
    name = "name",
    type = "type",
    host = "host",
    connect_code = "",
    connection_object = "",
    icon = "/usr/home/edgar/R/x86_64-pc-linux-gnu-library/3.6/connections/images/package-icon.png",
    disconnect = function() connection_close(my_conn, "host", "type"),
    preview_object = function() {},
    catalogs = list(
      name = "Database",
      type = "catalog",
      schemas = list(
        name = "Schema",
        type = "schema",
        tables = list(
          list(
            name = "table1",
            type = "table",
            fields = list(
              name = "field1",
              type = "chr")
            ),
          list(
            code = list(as.list(data.frame(name = "view1", type = "view", stringsAsFactors = FALSE)))  
          )
        ))
    ))

conn_list <- connection_contract(my_conn)
connection_view(conn_list)
```

<img src = "man/figures/custom-1.png" width = "400px"> <br/>
