---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# connections

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/edgararuiz/connections.svg?branch=master)](https://travis-ci.com/edgararuiz/connections)
[![Codecov test coverage](https://codecov.io/gh/edgararuiz/connections/branch/master/graph/badge.svg)](https://codecov.io/gh/edgararuiz/connections?branch=master)
<!-- badges: end -->

Provides a generic implementation of the [RStudio Connection Contract](https://rstudio.github.io/rstudio-extensions/connections-contract.html) to make it easier for database connections, and other type of connections, opened via R packages to take advantage of the Connections Pane inside the RStudio IDE.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("edgararuiz/connections")
```

## Usage

### Basic

```{r}
library(connections)
library(RSQLite)
library(DBI)

con <- connection_open(SQLite(), path = ":dbname")
```

```{r}
dbWriteTable(con, "db_mtcars", mtcars)
```

```{r}
connection_close(con)
```

### Not integrated

```{r}
con <- dbConnect(SQLite(), path = ":dbname")
connection_view(con)
```

```{r}
dbWriteTable(con, "db_mtcars", mtcars)
```

```{r}
connection_update(con)
```

```{r}
connection_close(con)
```

```{r}
con <- dbConnect(SQLite(), path = ":dbname")

connection_view(
  con, 
  host = "my_host", 
  name = "my_name",
  connection_code = "library(connections)\ndbConnect(...)"
  )
```

```{r}
connection_close(con)
```

## `DBI` packages examples

### `bigrquery`

```{r}
library(DBI)
library(bigrquery)

con <- dbConnect(
  bigquery(),
  project = "bigquery-public-data",
  dataset = "austin_311",
  billing = "rstudio-bigquery-event",
  use_legacy_sql = FALSE
)
```

```{r}
connection_view(con)
```






