---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
if (dir.exists("~/pins")) unlink("~/pins", recursive = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

add_figure <- function(path, width = 400) {
  char_html <- paste0("<img src='", path, "' width ='", width, "px'/><br/>")
  htmltools::HTML(char_html)
}

toc <- function() {
  re <- readLines("README.Rmd")
  has_title <- as.logical(lapply(re, function(x) substr(x, 1, 2) == "##"))
  only_titles <- re[has_title]
  titles <- trimws(gsub("#", "", only_titles))
  links <- trimws(gsub("`", "", titles))
  links <- tolower(links)
  links <- trimws(gsub(" ", "-", links))
  links <- trimws(gsub(",", "", links))
  toc_list <- lapply(
    seq_along(titles),
    function(x) {
      pad <- ifelse(substr(only_titles[x], 1, 3) == "###", "    - ", "  - ")
      paste0(pad, "[", titles[x], "](#", links[x], ")")
    }
  )
  toc_full <- paste(toc_list, collapse = "\n")
  cat(toc_full)
}
```

# connections

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/rstudio/connections/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rstudio/connections/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/rstudio/connections/branch/main/graph/badge.svg)](https://app.codecov.io/gh/rstudio/connections?branch=main)
<!-- badges: end -->

  - [Installation](#installation)
  - [Functions](#functions)
  - [Uploading and referencing tables with `dplyr`](#uploading-and-referencing-tables-with-dplyr)
  - [`pins`](#pins)
    - [Pin a database connection](#pin-a-database-connection)
    - [Pin a `dplyr` database query](#pin-a-dplyr-database-query)
    - [Full `pins` example](#full-pins-example)
  - [Back-end examples](#back-end-examples)
    - [BigQuery, via `bigrquery`](#bigquery-via-bigrquery)
    - [PostgreSQL, via `RPostgres`](#postgresql-via-rpostgres)
  - [`DBI` connections](#dbi-connections)

The main goal of `connections` is to integrate `DBI`-compliant packages with the
RStudio IDE's [Connection
Pane](https://db.rstudio.com/rstudio/connections/).  Packages such as 
[RPostgres](https://github.com/r-dbi/RPostgres), [RSQLite](https://github.com/r-dbi/RSQLite), 
[RMariaDB](https://github.com/r-dbi/RMariaDB) and
[bigrquery](https://github.com/r-dbi/bigrquery) connect R to those databases, but 
do not provide a direct integration with the Connections Pane. `connections` 
reads the configuration of the connection and creates the integration with RStudio.

A second goal is to provide integration with the [pins](https://pins.rstudio.com/)
package.  The `connections` package allows you to pin database connections and 
[dplyr](https://dplyr.tidyverse.org/) table objects. 

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("rstudio/connections")
```
```{r setup, include = FALSE}
library(connections)
library(RSQLite)
library(dplyr)
library(pins)

if (file.exists("local.sqlite")) unlink("local.sqlite")
```

## Functions

The two main functions added by `connections` are: 

- `connection_open()` - Opens the database connection. Use instead of 
`dbConnect()`, but use the exact same arguments. It also automatically starts 
the Connections pane. 
- `connection_close()` - Closes the database connection.

```{r}
library(connections)
library(RSQLite)

con <- connection_open(SQLite(), "local.sqlite")
```


`r add_figure("man/figures/connection-1.png")`

The connection can now be closed by using the appropriate button in the 
Connections pane, or by using `connection_close()`

```{r}
connection_close(con)
```

`r add_figure("man/figures/connection-2.png")`

The connection code is parsed when connecting to the database, and it is 
visible once the connection is closed.

## Uploading and referencing tables with `dplyr`

`connections` integrates with `dplyr` by supporting the following two functions: 

- `tbl()` - To create a pointer to a table or view within the database.
- `copy_to()` - To copy data from the R session to the database.

The version of `copy_to()` inside `connections` automatically updates the 
Connections pane, so the new table automatically shows up.

```{r}
con <- connection_open(SQLite(), "local.sqlite")

copy_to(con, mtcars, temporary = FALSE, overwrite = TRUE)
```

To use an existing table inside the database use `tbl()`.

```{r}
db_mtcars <- tbl(con, "mtcars")
```

`r add_figure("man/figures/pane-1.png")`

The `tbl()` function opens the rest of the already available `dplyr` 
database integration.

```{r}
db_mtcars %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))
```

## `pins`

The `connections` package integrates with `pins`.  It adds the ability to "pin"
database connections and queries. It follows the same approach as the
[vetiver](https://rstudio.github.io/vetiver-r/reference/vetiver_pin_write.html) 
package. `connections` now has two new functions:

- `connection_pin_write()`
- `connection_pin_read()`

### Pin a database connection

The `connection_pin_write()` function does **not** save the R object. It records
the code necessary to recreate the connection.

```{r}
library(pins)
board <- board_folder("~/pins")

connection_pin_write(board, con, name = "my_conn")
```

`r add_figure("man/figures/pins-1.png")`

If you wish to see the code that `connections` will use when recreating the
conneciton from the pin, you can use `connection_code()`:

```{r}
connection_code(con)
```

`connection_pin_read()` will replay the exact same code used to initially 
connect to the database.  Assign the output to a variable, such as
`con1`.  The variable will work just like any connection variable.

```{r}
con1 <- connection_pin_read(board, "my_conn")
```

The `con1` variable is now a regular database connection variable.

```{r}
db_mtcars <- tbl(con1, "mtcars") %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))

db_mtcars
```

### Pin a `dplyr` database query

When `dplyr` works with database data, the resulting query is not executed until 
the data is explicitly collected into R, or when printing the top results to the 
R Console.  The `pin` records two things:

- The `dplyr` R object that contains all of the transformations.  **It does not
save the actual results**.

- The necessary information to recreate the database connection. This is to make
sure that the data is being retrieved from the original database connection.

```{r}
connection_pin_write(board, db_mtcars, name = "avg_mpg")
```

`r add_figure("man/figures/pins-2.png")`

`connection_pin_read()` will connect to the database, and return the `dplyr` object. 
Without assigning it to a variable, the pin will immediately print the results of the 
database.  Those results are being processed at the time `connection_pin_read()` runs. 

```{r}
connection_pin_read(board, "avg_mpg")
```

### Full `pins` example

The way `pins` integrates with databases, via the `connections` package, allows 
to open the connection from a pin, and pipe all of the subsequent code into a
new pin.  Afterwards, that pin can be used to collect or to continue using the 
`dplyr` object.

```{r}
board <- board_folder("~/pins")

con <- connection_pin_read(board, "my_conn")

tbl_summary <- con %>%
  tbl("mtcars") %>%
  group_by(cyl) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))


connection_pin_write(board, tbl_summary, name = "cyl_mpg")

connection_close(con)

connection_pin_read(board, "cyl_mpg")
```

`r add_figure("man/figures/pins-3.png")`

## Back-end examples

There are a couple of examples of how the Connections pane will look when 
opening the connection via `connections`.

### BigQuery, via `bigrquery`

```{r, eval = FALSE}
library(connections)
library(bigrquery)

con <- connection_open(
  bigquery(),
  project = "bigquery-public-data",
  dataset = "austin_311",
  billing = "my_project_billing",
  use_legacy_sql = FALSE
)
```

`r add_figure("man/figures/bigquery-1.png")`

```{r, eval = FALSE}
connection_close(con)
```

`r add_figure("man/figures/bigquery-2.png")`


###  PostgreSQL, via `RPostgres`

```{r, eval = FALSE}
library(connections)
library(RPostgres)
con <- connection_open(
  Postgres(),
  host = "localhost",
  dbname = "datawarehouse",
  user = "[user id]",
  password = "[password]",
  bigint = "integer",
  port = "5432"
)
```

`r add_figure("man/figures/postgres-1.png")`


## `DBI` connections

It is possible to integrate `DBI` connections not opened via `connection_open()`.
To do that, use `connection_view()` and pass it the variable containing the 
existing database connection.  

```{r, eval = FALSE}
library(DBI)

con <- dbConnect(RSQLite::SQLite(), ":memory:")

connection_view(con)
```

`r add_figure("man/figures/dbi-1.png", 200)`


Changes to the database will not automatically load in the Connections pane. 
The `connection_update()` function will refresh the pane with the latest.

```{r, eval = FALSE}
dbWriteTable(con, "mtcars", mtcars)

connection_update(con)
```

`r add_figure("man/figures/dbi-2.png", 300)`

```{r, eval = FALSE}
connection_close(con)
```

`r add_figure("man/figures/dbi-3.png", 300)`
