---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

add_figure <- function(path, width = "300px") {
  char_html <- paste0("<img src='", path,"' style='display: block; margin-left: auto; margin-right: auto; width: ", width,";'/><br/>")
  htmltools::HTML(char_html)
}

toc <- function() {
  re <- readLines("README.Rmd")
  has_title <- as.logical(lapply(re, function(x) substr(x, 1, 2) == "##"))
  only_titles <- re[has_title]
  titles <- trimws(gsub("#", "", only_titles))
  links <- trimws(gsub("`", "", titles))
  links <- tolower(links)
  links <- trimws(gsub(" ", "-", links))
  toc_list <- lapply(
    seq_along(titles),
    function(x) {
      if(substr(only_titles[x], 1, 3) == "###") {
        paste0("<ul><li><a href=#", links[x], ">", titles[x], "</a></li></ul>")  
      } else {
        paste0("<li><a href=#", links[x], ">", titles[x], "</a></li>")  
      }
    }
  )
  toc_full <- paste(toc_list, collapse = "") 
  paste0("<ul>", toc_full, "</ul>")
}
```

# connections

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/edgararuiz/connections.svg?branch=master)](https://travis-ci.com/edgararuiz/connections)
[![Codecov test coverage](https://codecov.io/gh/edgararuiz/connections/branch/master/graph/badge.svg)](https://codecov.io/gh/edgararuiz/connections?branch=master)
<!-- badges: end -->

The main goal of `connections` is to integrate `DBI`-compliant packages with the RStudio IDE's [Connection
Pane](https://db.rstudio.com/rstudio/connections/).  Packages such as [RPostgres](https://github.com/r-dbi/RPostgres), [RSQLite](https://github.com/r-dbi/RSQLite), [RMariaDB](https://github.com/r-dbi/RMariaDB) and [bigrquery](https://github.com/r-dbi/bigrquery) connect R to those databases, but do not provide a direct integration with the Connections Pane. `connections` reads the configuration of the connection and creates the integration with RStudio.

A second goal is to provide integration with the [pins](https://rstudio.github.io/pins/) package.  The `connections` package allows you to pin database connections and [dplyr](https://dplyr.tidyverse.org/) table objects. 

A third goal of `connections` is to provide a simpler API to the way RStudio Connections pane integrates with R.  This is meant for use by advanced R developers who wish to create custom connection configurations, or for `DBI`-compliant package developers who wish to directly integrate the Connections pane with their package. 

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("edgararuiz/connections")
```
```{r setup, include = FALSE}
library(connections)
library(RSQLite)
library(dplyr)
library(pins)

board_register_local(cache = "~/pins")
pin_remove("my_conn", "local")
pin_remove("avg_mpg", "local")
pin_remove("cyl_mpg", "local")

if(file.exists("local.sqlite")) unlink("local.sqlite")
```

## Using

The two main functions added by `connections` are: 

- `connection_open()` - Opens the database connection. Use instead of `dbConnect()`, but use the exact same arguments. It also automatically starts the Connections pane. 
- `connection_close()` - Closes the database connection.

```{r}
library(connections)
library(RSQLite)

con <- connection_open(SQLite(), "local.sqlite")
```


`r add_figure("man/figures/connection-1.png")`

The connection can now be closed by using the appropriate button in the Connections pane, or by using `connection_close()`

```{r}
connection_close(con)
```

`r add_figure("man/figures/connection-2.png")`

The connection code is parsed when connecting to the database, and it is visible once the connection is closed.

### `dplyr`

`connections` integrates with `dplyr` by supporting the following two functions: 

- `tbl()` - To create a pointer to a table or view within the database.
- `copy_to()` - To copy data from the R session to the database.

The version of `copy_to()` inside `connections` automatically updates the Connections pane, so the new table automatically shows up.

```{r}
con <- connection_open(SQLite(), "local.sqlite")

copy_to(con, mtcars, temporary = FALSE, overwrite = TRUE)
```

To use an existing table inside the database use `tbl()`.

```{r}
db_mtcars <- tbl(con, "mtcars")
```

`r add_figure("man/figures/pane-1.png")`

The `tbl()` function opens the rest of the already available `dplyr` database integration.

```{r}
db_mtcars %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))
```

## `pins`

The `connections` package integrates with `pins`.  It enables the ability to save and retrieve connections and queries. 

```{r}
library(pins)
board_register_local(cache = "~/pins")
```

### Pin a connection

Use the same `pin()` command to save a database connection.  Under the hood, `connections` saves the **necessary information to recreate the connection code, not the actual connection R object**.  

```{r}
pin(con, "my_conn", board = "local")
```

`r add_figure("man/figures/pins-1.png")`

Use `pin_get()` to re-open the connection.  In effect, `pin_get()` will replay the exact same code used to initially connect to the database.  This means that `connection_open()` is already called for you, so the Connections pane should automatically start up.

```{r}
con <- pin_get("my_conn", board = "local")
```

Assign the output of `pin_get()` to a variable, such as `con`.  The variable will work just like any connection variable.

```{r}
db_mtcars <- tbl(con, "mtcars") %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))

db_mtcars
```

### Pin a `dplyr` database query

When `dplyr` works with database data, the resulting query is not executed until the data is explicitly collected into R, or when printing the top results to the R Console.  The `pin` records two things:

- The `dplyr` R object that contains all of the transformations.  **It does not save the actual results**.
- The necessary information to recreate the database connection. This is to make sure that the data is being retrieved from the original database connection.

```{r}
pin(db_mtcars, "avg_mpg", board = "local")
```

`r add_figure("man/figures/pins-2.png")`

`pin_get()` will connect to the database, and return the `dplyr` object.  Without assigning it to a variable, the pin will immediately print the results of the database.  Those results are being processed at the time `pin_get()` runs. 

```{r}
pin_get("avg_mpg", board = "local")
```

### Full `pins` example

The way `pins` integrates with databases, via the `connections` package, allows to open the connection from a pin, and pipe all of the subsequent code into a new pin.  Afterwards, that pin can be used to collect or to continue using the `dplyr` object.

```{r}
pin_get("my_conn", board = "local") %>%
  tbl("mtcars") %>%
  group_by(cyl) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE)) %>%
  pin("cyl_mpg", board = "local")

pin_get("cyl_mpg", board = "local")
```

`r add_figure("man/figures/pins-3.png")`

