---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

toc <- function() {
  re <- readLines("README.Rmd")
  has_title <- as.logical(lapply(re, function(x) substr(x, 1, 2) == "##"))
  only_titles <- re[has_title]
  titles <- trimws(gsub("#", "", only_titles))
  links <- trimws(gsub("`", "", titles))
  links <- tolower(links)
  links <- trimws(gsub(" ", "-", links))
  toc_list <- lapply(
    seq_along(titles),
    function(x) {
      if(substr(only_titles[x], 1, 3) == "###") {
        paste0("<ul><li><a href=#", links[x], ">", titles[x], "</a></li></ul>")  
      } else {
        paste0("<li><a href=#", links[x], ">", titles[x], "</a></li>")  
      }
    }
  )
  toc_full <- paste(toc_list, collapse = "") 
  paste0("<ul>", toc_full, "</ul>")
}
```

# connections

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/edgararuiz/connections.svg?branch=master)](https://travis-ci.com/edgararuiz/connections)
[![Codecov test coverage](https://codecov.io/gh/edgararuiz/connections/branch/master/graph/badge.svg)](https://codecov.io/gh/edgararuiz/connections?branch=master)
<!-- badges: end -->

The main goal of `connections` is to integrate `DBI`-compliant packages with the RStudio IDE's [Connection
Pane](https://db.rstudio.com/rstudio/connections/).  Packages such as [RPostgres](https://github.com/r-dbi/RPostgres), [RSQLite](https://github.com/r-dbi/RSQLite), [RMariaDB](https://github.com/r-dbi/RMariaDB) and [bigrquery](https://github.com/r-dbi/bigrquery) connect R to those databases, but do not provide a direct integration with the Connections Pane. `connections` reads the configuration of the connection and creates the integration with RStudio.

A second goal is to provide integration with the [pins](https://rstudio.github.io/pins/) package.  The `connections` package allows you to pin database connections and [dplyr](https://dplyr.tidyverse.org/) table objects. 

A third goal of `connections` is to provide a simpler API to the way RStudio Connections pane integrates with R.  This is meant for use by advanced R developers who wish to create custom connection configurations, or for `DBI`-compliant package developers who wish to directly integrate the Connections pane with their package. 

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("edgararuiz/connections")
```
```{r setup, include = FALSE}
library(connections)
library(RSQLite)
library(dplyr)
library(pins)

board_register_local(cache = "~/pins")
pin_remove("my_conn", "local")
pin_remove("avg_mpg", "local")
pin_remove("cyl_avg", "local")

if(file.exists("local.sqlite")) unlink("local.sqlite")
```

## Using

The two main functions added by `connections` are: 

- `connection_open()` - Opens the database connection. Use instead of `dbConnect()`, but use the exact same arguments. It also automatically starts the Connections pane. 
- `connection_close()` - Closes the database connection.

```{r}
library(connections)
library(RSQLite)

con <- connection_open(SQLite(), "local.sqlite")
```

```{r}
connection_close(con)
```

### Integration with `dplyr`

```{r}
con <- connection_open(SQLite(), "local.sqlite")
```

```{r}
copy_to(con, mtcars, temporary = FALSE, overwrite = TRUE)
```

```{r}
db_mtcars <- tbl(con, "mtcars")
```


```{r}
db_mtcars %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))
```

```{r}
connection_close(con)
```

## `pins`

```{r}
library(pins)
board_register_local(cache = "~/pins")
```

### Pin a connection

```{r}
pin(con, "my_conn", board = "local")
```

```{r}
connection_close(con)
```

```{r}
con <- pin_get("my_conn", board = "local")
```

```{r}
tbl(con, "mtcars")
```

### Pin a `dplyr` database query

```{r}
x <- tbl(con, "mtcars") %>%
  group_by(am) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE))

x
```


```{r}
pin(x, "avg_mpg", board = "local")
```

```{r}
connection_close(con)
```

```{r}
pin_get("avg_mpg", board = "local")
```

### Full `pins` example

```{r}
pin_get("my_conn", board = "local") %>%
  tbl("mtcars") %>%
  group_by(cyl) %>%
  summarise(avg_mpg = mean(mpg, na.rm = TRUE)) %>%
  pin("cyl_mpg", board = "local")
```

```{r}
pin_get("cyl_mpg", board = "local")
```
